Artificial Intelligence for the Automated Synthesis and Validation of Programs

### 2018-10-03 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updates to the conselleria project proposal

\section{Expected benefits of the research}
\label{subsec:beneficios}
{\em AI Planning} has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}. Promising research opportunities come from the application of {\em AI Planning} to {\em program synthesis} given that, {\em program synthesis} with test cases, can be seen as the {\em program testing} dual. In fact, our current work on {\em program synthesis} with AIP already produced several {\bf publications at the top international conferences and journals on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16,segovia:FSC:JAIR2018,segovia:programs:AIJ19} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.

The main benefit of this project is to {\bf provide new insights into the current understanding of how AI can assist programmers in the software development}.  Automated program synthesis promises to produce huge cost savings in comparison to conventional software development and to increase the quality, reliability and modifiability of programs.  In more detail, the expected benefits for this particular research project are four-fold:
\begin{enumerate}
\item An empirical {\bf study on the performance of the state-of-the-art AI planners for the {\em Synthesis and Validation of Programs}}. Research in AI algorithms is too often tested with laboratory problems and AIP is not an exception. Most of the new planning algorithms are only tested within the benchmarks of the International Planing Competition~\cite{vallati:IPC:AI15}. This project will help to meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks. 
\item A new {\bf evaluation methodology for the {\em Synthesis and Validation of Programs}}. The application of the exiting {\em AI Planning} technology to program synthesis can provide new evaluation metrics that assess how well a program covers a set of {\em input-output} test cases.  
\item The {\bf development of open software and open benchmarks for the {\em Synthesis and Validation of Programs}}. We strongly belief that reproducibility and open knowledge are essential to the advance of the research on computer science. With this regard, we plan to develop a {\em github} repository where we make available the developed source code and benchmarks.
\item {\bf International dissemination of the obtained scientific results}. The scientific results obtained during the development of the project will be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).
\end{enumerate}

\subsubsection{Gender sensitive programming}

A programming process is {\em gender sensitive} when the gender dimension is systematically integrated into everystep of the process~\cite{leduc2009guidelines}. From defining the problem, to identifying potential solutions, in the methodology and approach to implementing the project, in stakeholders analysis and the choice of partners, in defining the objective, outcomes, outputs, and activities, in the composition of the implementation and management team, in budgeting, in the monitoring and evaluation process, and in policy dialogue.

When designing a programme, it is important to take this dimension into account. However it is not easy with the current lack of representation of women, minorities in the IT work forces~\cite{arnold2001global}. Futher, statistically the roles of women in the IT sector have  greater representation  maintaining  legacy  systems  (using languages such as) rather than engineering new systems (using languages such as Java or C++~\cite{Dattero:2004:PLG:962081.962087}.

In this propoject we propose to increase the automation of the programming process (which arguably could improve the chances of developing programs less gender biased). However human are still required in the loop for deffining the set of input-output test cases that define the program behaviour.

>>>>

\section{Expected benefits of the research}
\label{subsec:beneficios}
{\em AI Planning} has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}. Promising research opportunities come from the application of {\em AI Planning} to {\em program synthesis} given that, {\em program synthesis} with test cases, can be seen as the {\em program testing} dual. In fact, our current work on {\em program synthesis} with AIP already produced several {\bf publications at top international conferences and journals on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16,javi-icaps17,segovia:FSC:JAIR2018,segovia:programs:AIJ19} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.

The main benefit of this project is to provide {\bf new insights into the current understanding of how AI can assist programmers} in the software development.  Automated program synthesis promises to produce huge cost savings in comparison to conventional software development and to increase the quality, reliability and modifiability of programs.  In more detail, the expected benefits for this particular research project are four-fold:
\begin{enumerate}
\item An empirical {\bf study on the performance of the state-of-the-art AI planners for the {\em Synthesis and Validation of Programs}}. Research in AI algorithms is too often tested with laboratory problems and AIP is not an exception. Most of the new planning algorithms are only tested within the benchmarks of the International Planing Competition~\cite{vallati:IPC:AI15}. This project will help to meet the computational and expressiveness limits of the off-the-shelf AI planners when addressing real-world programming tasks. 
\item A new {\bf evaluation methodology for the {\em Synthesis and Validation of Programs}}. The application of the exiting {\em AI Planning} technology to program synthesis can provide new evaluation metrics that assess how well a program covers a set of {\em input-output} test cases.  
\item The development of {\bf open software and benchmarks for the {\em Synthesis and Validation of Programs}}. We strongly belief that reproducibility and open knowledge are essential to the advance of the research on computer science. With this regard, we plan to develop a {\em github} repository where we make available the developed source code and benchmarks.
\item {\bf International dissemination of the obtained scientific results}. The scientific results obtained during the development of the project will be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).
\end{enumerate}

\subsubsection*{Automated Synthesis and Validation of Programs: A gender perspective}
Programs are {\em gender sensitive} when the gender dimension is systematically integrated into every step of the programming process~\cite{leduc2009guidelines}. From defining the problem, to identifying potential solutions, in the methodology and the final approach followed to eventually implement the program.

The consideration of the gender dimension at all the steps that are comprised in a programming process is nowadays a chimera. Specially, given the current lack of representation of women in the IT work forces~\cite{arnold2001global}.  In this project we propose to increase the automation of the programming process, which arguably, could improve the chances of developing programs that result less gender-biased.

In practice, our approach for automated program synthesis still requires humans in the loop (e.g. to specify the set of test cases that determine the program behavior). Therefore, even programs automatically synthesized can result gender-biased provided a gender-biased selection of test cases. This concern becomes more evident when we consider that statistically, the current roles of women in the IT sector have greater representation in maintaining {\em legacy systems} rather than in the engineering of {\em cutting-edge} systems~\cite{Dattero:2004:PLG:962081.962087}, like Artificial Intelligence systems.
